---
title: "Cause of Death"
author: "Tymur Tkachenko, Rohen Giralt, Kaeshev Alapati"
date: "2022-11-16"
output: html_document
---
<!DOCTYPE html>
<html>
    <head>
        <link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Bree+Serif&display=swap" rel="stylesheet">
<title>Online HTML Editor</title>
</head>
<body>
<h1>Online HTML Editor</h1>
<div>This is real time online HTML Editor</div>

<style>
  .section-header {
      display: grid;
  }
  .section-header-box {
      
      grid-column: 1;
      grid-row: 1;
    background-color: #b1d235;
    padding: .1em;
    font-size: 1em;
    align: center;
    margin: 3em 8em;
    text-align: center;
    font-family: "Bree Serif";
    color: white;
  }
  
  .section-header-line {
    background-color: black;
    padding: .1em;
    margin: 3em 8em;
    grid-column: 1;
      grid-row: 1;
      color:red;
  }
  
</style>
<div class="section-header">
<div class="section-header-box">
    <h1>Section 1</h1>
</div>

</div class="section-header-line">
    <h1>asdf</h1>
    
</div>
</div>

</body>
</html>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
```

Reading in the data
```{r message=FALSE, warning=FALSE}
data <- read_csv("data/deaths.csv")
```
Exploring the data set
```{r}
# knitr::kable(data)
head(data)
```
Cleaning up the data
```{r}
data = data %>% rename(Cause = `Cause of Death`)
data$Cause = as.factor(data$Cause)
data$State = as.factor(data$State)
summary(data)
```
## Question 10
```{r}
head(data %>% group_by(Cause) %>% summarize(total = sum(Deaths)/22) %>% arrange(desc(total)), n=10)
```

## Question 11
```{r}
head(data %>% group_by(Cause) %>% summarize(total = sum(Deaths)) %>% arrange(desc(total)), n=10)
```

None of the leading causes of death change when comparing the average over 22 years and the total. Since each number is changed equally to take the average, the order does not change.

## Question 12
```{r}
deaths = sum(data %>% filter(Year == "2015" & Cause == "Influenza and pneumonia (J09-J18)") %>% select(Deaths))
totalDeaths = sum(data %>% filter(Year == "2015") %>% select(Deaths))
prob = deaths/totalDeaths
```
The probability of having Influenze and pneumonia as the cause of death in 2015 is `r prob`.

## Question 13
```{r}
data %>% group_by(Year) %>% filter(State == "New York" & Cause == "Chronic lower respiratory diseases (J40-J47)") %>% select(Deaths) %>% arrange(desc(Deaths))
```
The most likely year of death is 1999
```{r}
deaths = sum(data %>% filter(Year == "1999" & State == "New York" & Cause == "Chronic lower respiratory diseases (J40-J47)") %>% select(Deaths))
totalDeaths = sum(data %>% filter(State == "New York" & Cause == "Chronic lower respiratory diseases (J40-J47)") %>% select(Deaths))
prob = deaths/totalDeaths
```
The probability of dying in 1999 if the cause of death is chronic lower respiratory diseases and the state is New York is `r prob`.

## Question 14
```{r}
deaths = sum(data %>% filter(Cause == "Malignant neoplasms (C00-C97)") %>% select(Deaths))
totalDeaths = sum(data %>% select(Deaths))
prob = deaths/totalDeaths
```
The probability of having died of cancer is `r prob`.

## Question 15
```{r}

```

## Question 16
```{r}
topDeaths = head(data %>% group_by(Cause) %>% summarize(total = sum(Deaths)) %>% arrange(desc(total)), n=6)
ggplot(data = topDeaths, aes(x="", y=total, fill=Cause)) +
  geom_col(width = 1, color="black") +
  coord_polar("y", start=0) +
  theme_void() +
  geom_text(aes(label = total), position = position_stack(vjust = 0.5))
```

## Question 17
```{r}
topDeaths = head(data %>% group_by(Cause) %>% summarize(total = sum(Deaths)) %>% arrange(desc(total)), n=10)

```

# Part 1

## 1. Describe in a few sentences two different groups, industries, sectors, or other areas of society that would experience a loss due to an increase in mortality rate in the United States.

## 2. Describe in no more than a few sentences an example of how insurance may be used to help mitigate or manage the losses you noted in question 1.

## 3. Describe in no more than a few sentences an example of how implementing new policies, procedures, or systems to help change behaviors of those involved to help mitigate or manage the losses you noted in this scenario.

## 4. Describe in no more than a few sentences an example of how those involved in the scenario may be able to help modify the possible outcomes to help mitigate or manage the losses you noted.

# Part 2

## 5. Based on the five categories outlined in the Actuarial Process Guide, what category (or categories) do you classify the attached dataset into? Provide a rationale for your reasoning.

## 6. Based on your answer to the previous question, identify one limitation there is on the types of questions or analysis that you can perform on the data?

## 7. Beyond the data provided in the attached spreadsheet, identify one other type of data or piece of information that could be valuable in analyzing the risks associated with evaluating cause of death and evaluation of impact of policy programs?

## 8. The researcher compiling this dataset compared select causes of death after pulling the raw data to the numbers from the last dataset published from the CDC for 1999-2017. The researcher compared the raw data to the last published dataset to check and see if any cleaning was needed. Use the table below to answer the following questions.

```{r echo=FALSE}
causes = c("Accidents (unintentional injuries)", 
           "All Causes",
           "Alzheimer's disease",
           "Cerebrovascular diseases",
           "Chronic lower respiratory diseases",
           "Diabetes Mellitus",
           "Diseases of heart",
           "Influenza and pneumonia",
           "Intentional self-harm (suicide)",
           "Malignant neoplasms",
           "Nephritis, nephrotic syndrome and nephrosis")
```
