---
title: "New Model"
author: "Rohen Giralt, Kaeshev Alapati, Tymur Tkachenko"
date: "2023-02-20"
output: html_document
---

## Libraries
```{r}
library(sqldf)
library(tidyverse)
library(tigerstats)
library(lattice)
library(lubridate)
library(geojsonR)
library(magrittr)
```

## Read in crimes data (may take a while!)
```{r}
crimes <- data.table::fread("data/Crimes_-_2001_to_Present.csv") # Faster than read_csv
```

```{r}

start_date <- "2020-1-1"
end_date <- "2022-12-31"

df_short <- crimes %>% select(Date, "Community Area", IUCR, "FBI Code", "Primary Type", Description) # Select desired columns
df_short <- df_short %>% rename(Community_Area = "Community Area", FBI_Code = "FBI Code", Primary_Type = "Primary Type", Secondary_Type = Description)

df_short %<>%
  mutate(Date = sub('.{12}$', '', df_short$Date)) %>% # idk kaeshev can u comment this
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
df_short <- df_short[df_short$Date > start_date, ] # Filter desired dates
df_short <- df_short[df_short$Date < end_date, ]

df_short %<>%
  mutate(Month = month(df_short$Date)) %>% # Add month column
  mutate(Year = year(df_short$Date)) %>% # Add year column
  mutate(DayMonth = mdy(sprintf("%d/01/%d",  Month, Year))) %>% # Add column based on month of each year (arbitrarily chosen to be first of each month)
  select(-Year) # Remove Year column

df_short <- as.data.frame(df_short) # Convert to data.frame (was previously data.table)
```

```{r}
df_short_inital <- df_short
```

## Read in temperature data
```{r}
temperature_data <- read_csv("data/Chicago area average temperature data.csv")
temperature_data %<>% pivot_longer(!Year, names_to = "month", values_to = "temp") %>%
  mutate(DayMonth = mdy(sprintf("%s/01/%d",  month, Year))) %>%
  select(temp, DayMonth)
```

## Combine datasets
```{r}
df_short %<>% merge(temperature_data, by="DayMonth")
rm(temperature_data)
```

# Read Community Areas number/name mapping data
```{r}
community_areas <- read_csv("data/CommAreas.csv")
community_areas %<>% select(AREA_NUMBE, COMMUNITY)
```
```{r}
school_data <- read_csv("data/Chicago_Public_Schools_-_School_Profile_Information_SY1617.csv")
school_locations <- read_csv("data/CPS_School_Locations_1617.csv")

school_locations %<>% select(COMMAREA, School_ID)
school_data %<>% merge(school_locations, "School_ID")
school_data$"COMMUNITY" <- school_data$COMMAREA
school_data %<>% merge(community_areas, "COMMUNITY")
school_data$Community_Area <- school_data$AREA_NUMBE

school_data %<>% select(Community_Area, Average_ACT_School, College_Enrollment_Rate_School, Student_Count_Low_Income, Student_Count_English_Learners, Student_Count_Special_Ed, Student_Count_Asian, Student_Count_Black, Student_Count_Hispanic, Student_Count_Multi, Student_Count_White, Student_Count_Total)
```
```{r}
Average_ACT_Community <-                    tapply(school_data$Average_ACT_School,              school_data$Community_Area, mean, na.rm=TRUE)
Average_College_Enrollment_Community <-     tapply(school_data$College_Enrollment_Rate_School,  school_data$Community_Area, mean, na.rm=TRUE)
Total_Asian_Count_Community <-              tapply(school_data$Student_Count_Asian,             school_data$Community_Area, sum, na.rm=TRUE)
Total_Black_Count_Community <-              tapply(school_data$Student_Count_Black,             school_data$Community_Area, sum, na.rm=TRUE)
Total_Hispanic_Count_Community <-           tapply(school_data$Student_Count_Hispanic,          school_data$Community_Area, sum, na.rm=TRUE)
Total_Multi_Count_Community <-              tapply(school_data$Student_Count_Multi,             school_data$Community_Area, sum, na.rm=TRUE)
Total_White_Count_Community <-              tapply(school_data$Student_Count_White,             school_data$Community_Area, sum, na.rm=TRUE)
Total_English_Learners_Count_Community <-   tapply(school_data$Student_Count_English_Learners,  school_data$Community_Area, sum, na.rm=TRUE)
Total_Low_Income_Count_Community <-         tapply(school_data$Student_Count_Low_Income,        school_data$Community_Area, sum, na.rm=TRUE)
Total_Student_Count_Community <-            tapply(school_data$Student_Count_Total,             school_data$Community_Area, sum, na.rm=TRUE)

Asian_Proportion_Community              <- Total_Asian_Count_Community              / Total_Student_Count_Community
Black_Proportion_Community              <- Total_Black_Count_Community              / Total_Student_Count_Community
Hispanic_Proportion_Community           <- Total_Hispanic_Count_Community           / Total_Student_Count_Community
Multi_Proportion_Community              <- Total_Multi_Count_Community              / Total_Student_Count_Community
White_Proportion_Community              <- Total_White_Count_Community              / Total_Student_Count_Community
English_Learners_Proportion_Community   <- Total_English_Learners_Count_Community   / Total_Student_Count_Community
Low_Income_Proportion_Community         <- Total_Low_Income_Count_Community         / Total_Student_Count_Community

condensed_school_data <- data.frame(Community_Area = 1:77, Average_ACT_Community, Average_College_Enrollment_Community, Asian_Proportion_Community, Black_Proportion_Community, Hispanic_Proportion_Community, Multi_Proportion_Community, White_Proportion_Community, English_Learners_Proportion_Community, Low_Income_Proportion_Community, Total_Student_Count_Community, Total_Asian_Count_Community, Total_Black_Count_Community, Total_Hispanic_Count_Community, Total_Multi_Count_Community, Total_White_Count_Community, Total_English_Learners_Count_Community, Total_Low_Income_Count_Community)
```

```{r}
df_short <- df_short %>% merge(condensed_school_data, "Community_Area")

rm(community_areas, school_locations, school_data)
rm(Average_ACT_Community, Average_College_Enrollment_Community, Total_Asian_Count_Community, Total_Low_Income_Count_Community, Total_Student_Count_Community, Asian_Proportion_Community, Black_Proportion_Community, Hispanic_Proportion_Community, Multi_Proportion_Community, English_Learners_Proportion_Community, Low_Income_Proportion_Community)
```

```{r}
life_expectancy <- read_csv("data/Public_Health_Statistics_-_Life_Expectancy_By_Community_Area_-_Historical.csv")
```

```{r}

housing <- read_csv("data/Affordable_Rental_Housing_Developments.csv") %>%
  select("Community Area Number", "Property Type", Units) %>%
  rename(Community_Area = "Community Area Number", Property_Type = "Property Type")
```
```{r}
# total units of each property type, by community
total_units_frame <- data.frame(tapply(housing$Units, list(housing$Property_Type, housing$Community_Area), sum))
total_units_frame %<>%
  mutate(Property_Type = rownames(total_units_frame)) %>%
  pivot_longer(cols = !Property_Type, names_to = "Community_Area", values_to = "Total_Units_Property_Community")
total_units_frame %<>%
  mutate(Community_Area = as.numeric(str_extract(total_units_frame$Community_Area, "\\d+"))) %>%
  drop_na()

missing_comm_areas <- data.frame(
  Community_Area = symdiff(1:77, housing$Community_Area),
  Property_Type = NA,
  Total_Units_Property_Community = rep(0, 77 - length(unique(housing$Community_Area)))
)

total_units_frame <- rbind(total_units_frame, missing_comm_areas, stringsAsFactors = FALSE)

housing %<>% merge(total_units_frame, c("Community_Area", "Property_Type"))

# x = data.frame(tapply(total_units_frame$Total_Units_Property_Community, list(total_units_frame$Community_Area, total_units_frame$Property_Type), sum))
# y = data.frame(tapply(housing$Units, list(housing$Property_Type, housing$Community_Area), sum))
# Total_Senior_Housing_Community =
# total_units_frame %>% mutate(
#   Total_Senior_Housing_Community =
# )
```

```{r}
df_short <- df_short %>% merge(housing, "Community_Area")
```


```{r}

# df_short <- df_short %>% mutate(
#   Senior_Housing_Count =               count(lapply(Property_Type, grepl, "Senior", fixed = TRUE)), # `grepl` is basically `contains`
#   ARO_Housing_Count =                  count(lapply(Property_Type, grepl, "ARO", fixed = TRUE)),
#   Intergenerational_Housing_Count =    count(lapply(Property_Type, grepl, "Inter-generational", fixed = TRUE)),
#   Multifamily_Housing_Count =          count(lapply(Property_Type, grepl, "Multifamily", fixed = TRUE)),
#   Supportive_Housing_Count =           count(lapply(Property_Type, grepl, "Supportive", fixed = TRUE)),
#   SRO_Housing_Count =                  count(lapply(Property_Type, grepl, "SRO", fixed = TRUE)),
#   Artist_Housing_Count =               count(lapply(Property_Type, grepl, "Artist", fixed = TRUE)),
# )
```

# costs
```{r}

crime_costs_prim_match <- read_csv("data/Crime Costs (Primary Type Matched).csv") # Crime costs for crimes indexed by their Primary Type
crime_costs_IUCR_match <- read_csv("data/Crime Costs (IUCR Matched).csv") # Crime costs for crimes indexed by IUCRs
crime_costs_IUCR_match$IUCR %<>% as.character()

df_short %<>%
  merge(crime_costs_prim_match, by="Primary_Type", all = TRUE) %>%
  merge(crime_costs_IUCR_match, by="IUCR", all = TRUE)

# Get rid of duplicates...
df_short %<>% mutate(
  Crime_Career_Cost =               coalesce(df_short$`Crime Victim Cost.x`,               df_short$`Crime Victim Cost.y`),
  Criminal_Justice_System_Cost =    coalesce(df_short$`Criminal Justice System Cost.x`,    df_short$`Criminal Justice System Cost.y`),
  Crime_Career_Cost =               coalesce(df_short$`Crime Career Cost.x`,               df_short$`Crime Career Cost.y`),
  Total_Tangible_Cost =             coalesce(df_short$`Total Tangible Cost.x`,             df_short$`Total Tangible Cost.y`),
  Pain_And_Suffering_Cost =         coalesce(df_short$`Pain and Suffering Cost.x`,         df_short$`Pain and Suffering Cost.y`),
  Risk_Of_Homicide_Cost =           coalesce(df_short$`Corrected Risk-of-Homicide Cost.x`, df_short$`Corrected Risk-of-Homicide Cost.y`),
  Total_Intangible_Cost =           coalesce(df_short$`Total Intangible Cost.x`,           df_short$`Total Intangible Cost.y`),
  Total_Total_Cost =                coalesce(df_short$`Total Total Cost.x`,                df_short$`Total Total Cost.y`),
  Total_Included_Cost =             coalesce(df_short$`Total Included Cost.x`,             df_short$`Total Included Cost.y`)
) %>%
  select(-c(`Crime Victim Cost.x`, `Crime Victim Cost.y`, `Criminal Justice System Cost.x`, `Criminal Justice System Cost.y`, `Crime Career Cost.x`, `Crime Career Cost.y`, `Total Tangible Cost.x`, `Total Tangible Cost.y`, `Pain and Suffering Cost.x`, `Pain and Suffering Cost.y`, `Corrected Risk-of-Homicide Cost.x`, `Corrected Risk-of-Homicide Cost.y`, `Total Intangible Cost.x`, `Total Intangible Cost.y`, `Total Total Cost.x`, `Total Total Cost.y`, `Total Included Cost.x`, `Total Included Cost.y`))

df_short <- df_short[!is.na(df_short$Total_Included_Cost),] # Remove ignored crimes (~15.54% of dataset)
```

# save
```{r}
save(df_short, file = "data/df_short.Rdata")
```

# Models
## Get data
```{r}
parsed_data <- df_short %>% count(temp, DayMonth, Community_Area, Average_ACT_Community, Average_College_Enrollment_Community, Asian_Proportion_Community, Black_Proportion_Community, Hispanic_Proportion_Community, Multi_Proportion_Community, White_Proportion_Community, English_Learners_Proportion_Community, Low_Income_Proportion_Community, Total_Asian_Count_Community, Total_Black_Count_Community, Total_Hispanic_Count_Community, Total_Multi_Count_Community, Total_White_Count_Community, Total_English_Learners_Count_Community, Total_Low_Income_Count_Community, Total_Student_Count_Community, Property_Type, Total_Units_Property_Community)
```
Dummies
```{r}
parsed_data <- parsed_data %>% mutate(
  housing_is_Senior =               as.vector(lapply(Property_Type, grepl, "Senior", fixed = TRUE), mode = "logical"), # `grepl` is basically `contains`
  housing_is_ARO =                  as.vector(lapply(Property_Type, grepl, "ARO", fixed = TRUE), mode = "logical"),
  housing_is_Intergenerational =    as.vector(lapply(Property_Type, grepl, "Inter-generational", fixed = TRUE), mode = "logical"),
  housing_is_Multifamily =          as.vector(lapply(Property_Type, grepl, "Multifamily", fixed = TRUE), mode = "logical"),
  housing_is_Supportive =           as.vector(lapply(Property_Type, grepl, "Supportive", fixed = TRUE), mode = "logical"),
  housing_is_SRO =                  as.vector(lapply(Property_Type, grepl, "SRO", fixed = TRUE), mode = "logical"),
  housing_is_Artist =               as.vector(lapply(Property_Type, grepl, "Artist", fixed = TRUE), mode = "logical"),
)
```

```
## Model
```{r}
# factors <- c(
#   parsed_data$Average_ACT_Community,
#   parsed_data$Average_College_Enrollment_Community,
#   parsed_data$Total_Asian_Count_Community,
#   parsed_data$Total_Black_Count_Community,
#   parsed_data$Total_Hispanic_Count_Community,
#   parsed_data$Total_Multi_Count_Community,
#   parsed_data$Total_White_Count_Community,
#   parsed_data$Total_Low_Income_Count_Community,
#   parsed_data$Total_English_Learners_Count_Community,
#   # Total_Student_Count_Community
#   parsed_data$temp
# )

model <- lm(
  n ~
  Average_ACT_Community +
  # Average_College_Enrollment_Community +
  Total_Asian_Count_Community +
  Total_Black_Count_Community +
  # Total_Hispanic_Count_Community +
  Total_Multi_Count_Community +
  Total_White_Count_Community +
  Total_Low_Income_Count_Community +
  Total_English_Learners_Count_Community +
  Total_Units_Property_Community * housing_is_Senior + # wait this is not the right way to do this
  Total_Units_Property_Community * housing_is_ARO +
  Total_Units_Property_Community * housing_is_Intergenerational +
  Total_Units_Property_Community * housing_is_Multifamily +
  Total_Units_Property_Community * housing_is_Supportive +
  Total_Units_Property_Community * housing_is_SRO +
  Total_Units_Property_Community * housing_is_Artist +
  # Total_Student_Count_Community +
  temp,
  data = parsed_data)
print(summary(model))

# model2 <- lm(
#   log(n) ~
#     log(Total_Black_Count_Community) +
#       log(Total_Student_Count_Community)
#       log(Total_Low_Income_Count_Community) + temp,
#   data = parsed_data
# )
print(summary(model2))

ggplot(parsed_data, aes(x = log(Total_Black_Count_Community), y = log(n))) +
  geom_point() +
  geom_smooth(method = "lm")


# print(summary(lm(log(n) ~ log(Total_Student_Count_Community), data = parsed_data)))
# print(summary(lm(n ~ log(Total_Student_Count_Community), data = parsed_data)))
# print(summary(lm(n ~ Total_Student_Count_Community, data = parsed_data)))

# ggplot(parsed_data, aes(x = log(Total_Student_Count_Community), y = log(n))) +
#         geom_point() +
#         geom_smooth(method = "lm")
#
# ggplot(parsed_data, aes(x = log(Total_Student_Count_Community), y = (n))) +
#         geom_point() +
#         geom_smooth(method = "lm")
#
# ggplot(parsed_data, aes(x = (Total_Student_Count_Community), y = (n))) +
#         geom_point() +
#         geom_smooth(method = "lm")


# for (factor in factor) {
#     ggplot(parsed_data, aes(x = factor, y = n)) +
#         geom_point() +
#         geom_smooth(method = "lm")
# }
```