---
title: "Beginning Crimes"
author: "Kaeshev ALapati"
date: "2023-02-02"
output: html_document
---

```{r}
library(sqldf)
library(tidyverse)
library(tigerstats)
library(lattice)
library(lubridate)
library(geojsonR)
library(magrittr)
#crime_data <- read.csv.sql("C:\\Users\\Kaeshev_Alapati\\Downloads\\Crimes_-_2001_to_Present.csv",
#                           sql = "select * from file where `Date` > 2021", eol = "\n")
# 
# con = file("C:\\Users\\Kaeshev_Alapati\\Downloads\\Crimes_-_2001_to_Present.csv", "r")
# csv = read.csv(con, nrows = 1000)
# close(con)
# df = data.frame(csv)
df <- read_csv("data/Crimes_-_2001_to_Present.csv")
housing_economics <- read_csv("C:/Users/Kaeshev_Alapati/Downloads/Census_Data_-_Selected_socioeconomic_indicators_in_Chicago__2008___2012.csv")
housing_economics = housing_economics %>% filter(!is.na(`Community Area Number`))
```

## Filtering and formatting crimes data
# note start_date, end_date vars
```{r}
start_date = "2020-1-1"
end_date = "2022-12-31"


df_short <- df %>%
  select(Date, "Community Area", "Primary Type", "IUCR")
df_short <- df_short %>%
  mutate(Date = sub('.{12}$', '', df_short$Date)) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
df_short <- df_short[df_short$Date > start_date, ]
df_short <- df_short[df_short$Date < end_date, ]
```

```{r}
df_short <- df_short %>% mutate(Wday = wday(df_short$Date))
df_short <- df_short %>% mutate(Month = month(df_short$Date))
df_short <- df_short %>% mutate(Year = year(df_short$Date))
head(mdy(sprintf("%d/01/%d",  df_short$Month, df_short$Year)))
df_short <- df_short %>% mutate(DayMonth = mdy(sprintf("%d/01/%d",  Month, Year)))
df_short <- df_short %>% mutate(MonthYear = format(Date, "%Y-%m"))
df_short <- df_short %>% mutate(condensed_crimes = `Primary Type`)
df_short <- df_short %>% mutate(Primary_Type = `Primary Type`)

# unique(df_short$`Primary Type`)
# df_categories = data.frame(category = c(replicate(15, "Costless Crime"), replicate(13, "Basic Crime"), replicate(3, "Extreme Crime")), 
#                            crime = c("CONCEALED CARRY LICENSE VIOLATION", "OBSCENITY", "PUBLIC INDECENCY", "NON-CRIMINAL", "DECEPTIVE PRACTICE", "GAMBLING", "PROSTITUTION", "NARCOTICS", "OTHER NARCOTIC VIOLATION", "INTERFERENCE WITH PUBLIC OFFICER", "INTIMIDATION", "LIQUOR LAW VIOLATION", "STALKING", "CRIMINAL TRESSPASS", "PUBLIC PEACE VIOLATION", "THEFT", "MOTOR VEHICLE THEFT", "ASSAULT", "ROBBERY", "BURGLARY", "BATTERY", "CRIMINAL DAMAGE", "KIDNAPPING", "OFFENSE INVOLVING CHILDREN", "SEX OFFENSE", "WEAPONS VIOLATION", "CRIMINAL SEXUAL ASSAULT", "OTHER OFFENSE", "ARSON", "HOMICIDE", "HUMAN TRAFFIKING"), 
#                            cost = c(replicate(15, 0), c(1000, 30000, 2000, 1000, 10000, 2000, 10000, 5000, 1000, 5000, 1000, 10000, 2000, 100000, 10000000, 1000000)))
# df_categories
# cat = c()
# cost = c()
# for (crime in df_short$`Primary Type`) {
#   cost = append(cost, df_categories$cost[df_categories$crime == crime])
#   cat = append(cat, df_categories$category[df_categories$crime == crime])
# }
# 
# df_short$cost <- df_categories$cost %>% group_by(df_categories$crime)
# 
# satest = c(1, 2, 3, 4)
# other = data.frame(letters = c("a", "b", "c", "d"), numbers = c(1, 2, 3, 4))
# 
# new_test = c()
# for (num in test) {
#   new_test = append(new_test, other$letters[num == other$numbers])
# }
# new_test
# 
# 
# 
# df_short <- df_short %>% mutate(category = replace(df_short, `Primary Type`, df_categories$category[df_categories$crime == `Primary Type`]))
# 
# df_short <- df_short %>% mutate(category = replace(df_categories$cost[df_categories$crime == crime]) df_categories$category[df_categories$crime == `Primary Type`], 
#                                 cost = df_categories$cost[df_categories$crime == `Primary Type`])
```
## Crime by DOTW bar
```{r}
# wday: Sun < Mon < Tue < Wed < Thu < Fri < Sat

days = 1:7
days_name = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")
crimes_by_day = c()
crimes_by_day_names = c()
for (day in days) {
  crimes_by_day = append(crimes_by_day, count(df_short$Wday == day))
}

crimes_by_day_frame = data.frame(day = days, crimes_by_day, day_name = days_name)

ggplot(data=crimes_by_day_frame, aes(x = day, y = crimes_by_day)) +
  geom_bar(stat="identity") + geom_text(aes(label = day_name), vjust = 1.5, colour = "white") + geom_text(aes(label = crimes_by_day), vjust = 3.5, colour = "white") + ggtitle(paste0("Crimes per day from ", start_date, " to ", end_date))
```
## Crimes by month bar
```{r}
months = 1:12
month_names = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
crimes_by_month = c()
for (month in months) {
  crimes_by_month = append(crimes_by_month, count(df_short[df_short$Date < "2023-1-1",]$Month == month))
}

# TODO: FIX THIS SHIT SOMEONE please!
crimes_by_month_frame = data.frame(month = months, crimes_by_month, month_name = month_names)
ggplot(data=crimes_by_month_frame, aes(x = months, y = crimes_by_month)) +
  geom_bar(stat="identity") + geom_text(aes(label = month_name), vjust = 1.5, colour = "white") + geom_text(aes(label = crimes_by_month), vjust = 3.5, colour = "white") + ggtitle(paste0("Crimes per month from ", start_date, " to ", end_date))
```
## Crimes by date timeseries
```{r}
dates <- seq(as.Date(start_date), as.Date(end_date), by="days")
crimes_by_date <- c()
for (date in dates) {
  crimes_by_date <- append(crimes_by_date, count(df_short$Date == date))
}
crimes_by_date_frame <- data.frame(date = dates, crimes_by_date)
head(crimes_by_date_frame)

ggplot(data = crimes_by_date_frame, aes(x = date, y = crimes_by_date)) + geom_line() + ggtitle(paste0("Crimes over time from ", start_date, " to ", end_date))

date_model <- lm(crimes_by_date ~ dates, crimes_by_date_frame)
print(summary(date_model))
```

## Crime proportions

```{r}
crime_counts = data.frame(count(df_short, df_short$`Primary Type`))
crime_counts %<>% rename(type_of_crime = df_short..Primary.Type., number_of_crimes = n) %>% mutate(prop = number_of_crimes / sum(number_of_crimes) *100) %>% arrange(desc(prop))
crime_counts_short = head(arrange(crime_counts, desc(prop)), 10)
crime_counts_short %<>% arrange(desc(prop))
other = data.frame(type_of_crime = "Other", number_of_crimes = sum(crime_counts$number_of_crimes[11:nrow(crime_counts)]), prop = sum(crime_counts$prop[11:nrow(crime_counts)]))
crime_counts_short = rbind(crime_counts_short, other)

ggplot(crime_counts_short, aes(x = "", y = prop, fill = reorder(type_of_crime, -prop))) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  theme_void() +
  scale_fill_brewer(palette="Spectral") + 
  labs(title = paste0("Crime Proportions From ", start_date, " to ", end_date)) + 
  labs(fill = "Type of Crime")
```


## Housing costs by community area
```{r}
qx_to_md <- function(quarters) {
  output = c()
  quarters_dates = data.frame(quarters = c("Q1", "Q2", "Q3", "Q4"), dates = c("2/15", "5/15", "8/15", "11/15"))
  for (input_quarter in quarters) {
    for (quarter in quarters_dates$quarters) {
      if (input_quarter == quarter) {
        output = append(output, quarters_dates$dates[quarters_dates$quarter == quarter])
      }
    }
  }
  return(output)
}

submarket_definitions = read_csv("C:/Users/Kaeshev_Alapati/OneDrive - Cary Academy/Staculus/submarket definitions.csv")
quarterly_values = read_csv("C:/Users/Kaeshev_Alapati/OneDrive - Cary Academy/Staculus/2022_q2_price_index_data.csv")
first_row = quarterly_values[1, ]
quarterly_values = quarterly_values[-1, ]
colnames(quarterly_values) = first_row
head(quarterly_values)
quarterly_values %<>% rename(date = YEARQ) %>% 
  mutate(year =  sub('.{2}$', '', date)) %>% 
  mutate(quarter = sub('.{4}', '', date)) %>% 
  mutate(quarter = qx_to_md(quarter)) %>% 
  mutate(date = as.Date(paste0(year, "/", quarter), format = "%Y/%m/%d")) %>% 
  select(-year, -quarter)

head(quarterly_values)

```
# costs
```{r}
crime_costs_prim_match <- read_csv("C:/Users/Kaeshev_Alapati/OneDrive - Cary Academy/Staculus/Crime Costs (Primary Type Matched).csv") # Crime costs for crimes indexed by their Primary Type
crime_costs_IUCR_match <- read_csv("C:/Users/Kaeshev_Alapati/OneDrive - Cary Academy/Staculus/Crime Costs (IUCR Matched).csv") # Crime costs for crimes indexed by IUCRs
crime_costs_IUCR_match$IUCR %<>% as.character()
df_short %<>%
  merge(crime_costs_prim_match, by="Primary_Type", all = TRUE) %>%
  merge(crime_costs_IUCR_match, by="IUCR", all = TRUE)
# Get rid of duplicates...
df_short %<>% mutate(
  Crime_Career_Cost =               coalesce(df_short$`Crime Victim Cost.x`,               df_short$`Crime Victim Cost.y`),
  Criminal_Justice_System_Cost =    coalesce(df_short$`Criminal Justice System Cost.x`,    df_short$`Criminal Justice System Cost.y`),
  Crime_Career_Cost =               coalesce(df_short$`Crime Career Cost.x`,               df_short$`Crime Career Cost.y`),
  Total_Tangible_Cost =             coalesce(df_short$`Total Tangible Cost.x`,             df_short$`Total Tangible Cost.y`),
  Pain_And_Suffering_Cost =         coalesce(df_short$`Pain and Suffering Cost.x`,         df_short$`Pain and Suffering Cost.y`),
  Risk_Of_Homicide_Cost =           coalesce(df_short$`Corrected Risk-of-Homicide Cost.x`, df_short$`Corrected Risk-of-Homicide Cost.y`),
  Total_Intangible_Cost =           coalesce(df_short$`Total Intangible Cost.x`,           df_short$`Total Intangible Cost.y`),
  Total_Total_Cost =                coalesce(df_short$`Total Total Cost.x`,                df_short$`Total Total Cost.y`),
  Total_Included_Cost =             coalesce(df_short$`Total Included Cost.x`,             df_short$`Total Included Cost.y`)
) %>%
  select(-c(`Crime Victim Cost.x`, `Crime Victim Cost.y`, `Criminal Justice System Cost.x`, `Criminal Justice System Cost.y`, `Crime Career Cost.x`, `Crime Career Cost.y`, `Total Tangible Cost.x`, `Total Tangible Cost.y`, `Pain and Suffering Cost.x`, `Pain and Suffering Cost.y`, `Corrected Risk-of-Homicide Cost.x`, `Corrected Risk-of-Homicide Cost.y`, `Total Intangible Cost.x`, `Total Intangible Cost.y`, `Total Total Cost.x`, `Total Total Cost.y`, `Total Included Cost.x`, `Total Included Cost.y`))
df_short <- df_short[!is.na(df_short$Total_Included_Cost),] # Remove ignored crimes (~15.54% of dataset)
```
## Cost Breakdown:
```{r}
crime_costs = data.frame(aggregate(Total_Included_Cost ~ Primary_Type , df_short, sum))
crime_costs %<>% mutate(prop = Total_Included_Cost / sum(Total_Included_Cost) * 100) %>% arrange(desc(prop))
crime_costs_short = head(crime_costs, 10)
other = data.frame(Primary_Type = "Other", Total_Included_Cost = sum(crime_costs$Total_Included_Cost[11:nrow(crime_costs)]), prop = sum(crime_costs$prop[11:nrow(crime_costs)]))
crime_costs_short %<>% rbind(other)
crime_costs_short

ggplot(crime_costs_short, aes(x = "", y = prop, fill = reorder(Primary_Type, -prop))) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  theme_void() +
  scale_fill_brewer(palette="Spectral") + 
  labs(title = paste0("Crime Cost Proportions From ", start_date, " to ", end_date)) + 
  labs(fill = "Type of Crime")

```
## Cost timescale by crimes
```{r}
dates <- format(seq(as.Date(start_date), as.Date(end_date), by="months"), "%Y-%m")
crime_types <- unique(df_short$Primary_Type)
comm_date = aggregate(Total_Included_Cost ~ Primary_Type + MonthYear, data = df_short, FUN=sum)
comm_date %<>% rename(date = MonthYear)
comm_date %<>% mutate(Primary_Type = as.factor(Primary_Type))
comm_date %<>% mutate(date = as.Date(paste0(date, "-01"), format = "%Y-%m-%d"))

ggplot(comm_date, aes(x=date, y=Total_Included_Cost, group=Primary_Type, color=reorder(Primary_Type, -Total_Included_Cost))) +
  geom_line() + 
  labs(title = paste0("The Costs of Crimes From ", start_date, " to ", end_date)) + 
  labs(color = "Type of Crime") +
  xlab("Date") +
  ylab("Cost to Chicago's Society ($)")

ggplot(comm_date, aes(x=date, y=log(Total_Included_Cost), group=Primary_Type, color=reorder(Primary_Type, -Total_Included_Cost))) +
  geom_line() + 
  labs(title = paste0("Logged Costs of Crimes From ", start_date, " to ", end_date)) + 
  labs(color = "Type of Crime") + 
  xlab("Date") + 
  ylab("Logged Cost to Chicago's Society ($)")
  
  
```
## Crime Cost House Cost model
```{r}


```

## number of crimes timescale
```{r}
crimes_over_time = count(df_short, Date, `Primary Type`)

types = unique(comm_date$`Primary_Type`)

crimes_over_time %<>% filter(is.element(`Primary Type`, types))

ggplot(crimes_over_time, aes(x=Date, y=n, group = `Primary Type`, color = reorder(`Primary Type`, -n))) + 
  geom_line()
```

## Crimes by community area
```{r}
community_areas = 1:77
dates <- format(seq(as.Date(start_date), as.Date(end_date), by="months"), "%Y-%m")

df_crimes_by_community_area_and_dates = data.frame(community_area = c(), date = c(), number_of_crimes = c())

for (community_area in community_areas) {
  for (date in dates) {
    new_row = data.frame(community_area = community_area, MonthYear = date, 
                         number_of_crimes = count(df_short$MonthYear == date & df_short$`Community Area` == community_area))
    
    df_crimes_by_community_area_and_dates = rbind(df_crimes_by_community_area_and_dates, new_row)
  }
}

df_crimes_by_community_area_and_dates = df_crimes_by_community_area_and_dates %>% mutate(MonthYear = as.Date(paste0(df_crimes_by_community_area_and_dates$MonthYear, "-01"), format=("%Y-%m-%d")), community_area = as.factor(community_area))

df_crimes_by_community_area_and_dates

ggplot(df_crimes_by_community_area_and_dates, aes(x = MonthYear, y = number_of_crimes, color = community_area), group = community_area) +
  geom_line() +
  ggtitle(paste0("Crimes over time by community area from ", start_date, " to ", end_date)) +
  labs(color = "Community Area") +
  xlab("Time") +
  ylab("Number of Crimes") +
  theme(legend.margin = margin(100, 1, 1, 1))
```

## Crimes by crime type
```{r}
categories = unique(df_short$`Primary Type`)

categories
df_crimes_by_type_timescale = data.frame(number_of_crimes = c(), date = c(), type_of_crime = c())

for (category in categories) {
  for (date in dates) {
    new_row = data.frame(type = type, 
                         date = date, 
                         number_of_crimes = count(date == df_short$MonthYear & 
                                                    type == df_short$`Primary Type`))
    
    df_crimes_by_type_timescale = rbind(df_crimes_by_type_timescale, new_row)
  }
}
head(df_crimes_by_type_timescale)

df_crimes_by_type_timescale = df_crimes_by_type_timescale %>% mutate(date = as.Date(paste0(date, "-01"), format=("%Y-%m-%d")), type = as.factor(type))

ggplot(df_crimes_by_type_timescale, aes(x = date, y = number_of_crimes, group = type, color = type)) + 
  geom_line() + 
  ggtitle(paste0("Crimes over time by crime type area from ", start_date, " to ", end_date)) +
  labs(color = "Crime Type") +
  xlab("Time") +
  ylab("Number of Crimes") +
  theme(legend.margin = margin(100, 1, 1, 1))
```

## Crimes by type and community area
```{r}
## rohen you're going to have to run this code my computer is too slow

df_crimes_by_type_and_community_area_timescale = c(date = c(), community_area = c(), type = c(), number_of_crimes = c())

for (date in dates) {
  for (type in types) {
    for (community_area in community_areas) {
      new_row = data.frame(date = date, type = type, 
                           community_area = community_area, 
                           number_of_crimes = count(date == df_short$MonthYear & 
                                                      type == df_short$`Primary Type` & 
                                                      community_area == df_short$`Community Area`))
      df_crimes_by_type_and_community_area_timescale = rbind(df_crimes_by_type_and_community_area_timescale, new_row)
    }
  }
}

head(df_crimes_by_type_and_community_area_timescale)

ggplot(df_crimes_by_type_and_community_area_timescale, aes(x = date, y = number_of_crimes, shape = community_area, color = type)) + 
  geom_line() + 
  ggtitle(paste0("Crimes over time by crime type and community area area from ", start_date, " to ", end_date)) +
  labs(color = "Crime Type", shape = "Community Area") +
  xlab("Time") +
  ylab("Number of Crimes") +
  theme(legend.margin = margin(100, 1, 1, 1))

```

## Crimes by location/socieconomics
```{r}
tail(housing_economics)
number_of_crimes_by_community_area = data.frame(community_area = c(), number_of_crimes = c(), cost_of_crimes = c())
for (community_area in community_areas) {
  {
    for (category in categories) {
      new_row = data.frame(community_area = community_area, number_of_crimes = count(community_area == df_short$`Community Area`))
      number_of_crimes_by_community_area= rbind(number_of_crimes_by_community_area, new_row)
    }
  }
}
data = data.frame(community_area = housing_economics$`Community Area Number`, crowded_housing = housing_economics$`PERCENT OF HOUSING CROWDED`)

model = lm(number_of_crimes ~ :`Community`, data = data)

```













## Useless <___>
```{r}
temperature_data <- read_csv("data/Chicago area average temperature data.csv")
temperature_data <- temperature_data %>% pivot_longer(!Year, names_to = "month", values_to = "temp")
temperature_data <- temperature_data %>% mutate(DayMonth = mdy(sprintf("%s/01/%d",  month, Year)))
```

```{r}
df_short = merge(df_short, temperature_data, by="DayMonth")
```

```{r}
temps = unique(df_short$temp)
crimes_by_temp = c()
for (temp in temps) {
  crimes_by_temp <- append(crimes_by_temp, count(df_short$temp == temp))
}
```
```{r}
crimes_by_temp_frame <- data.frame(temps, crimes_by_temp)
ggplot(crimes_by_temp_frame, aes(x = temps, y = crimes_by_temp)) +
    geom_point() +
    geom_smooth(method = "lm")
temp_model <- lm(crimes_by_temp ~ temps, crimes_by_temp_frame)
print(summary(temp_model))
```


```{r}
# community_areas_json <- FROM_GeoJson(url_file_string = "data/Boundaries - Community Areas (current).geojson")
# print(community_areas_json$features[[1]])
# print(typeof(community_areas_json$features[[1]]))
# print()
community_areas = read_csv("data/CommAreas.csv")
community_areas %>% select(AREA_NUMBE, COMMUNITY)
```
```{r}
school_data <- read_csv("data/Chicago_Public_Schools_-_School_Profile_Information_SY1617.csv")
school_locations <- read_csv("data/CPS_School_Locations_1617.csv")
school_locations <- school_locations %>% select(COMMAREA, School_ID)
school_data <- merge(school_data, school_locations, "School_ID")
school_data$COMMUNITY <- school_data$COMMAREA
school_data %>% select(-school_data$COMMAREA)
school_data = merge(school_data, community_areas, "COMMUNITY")
```
```{r}
community_areas = 1:77
asians_by_area = c()
crimes_by_area = c()
for (area in community_areas) {
  schools_in_area = school_data %>% filter(school_data$AREA_NUMBE == area)
  sum(schools_in_area$Student_Count_Asian)
  asians_by_area <- append(asians_by_area, sum(schools_in_area$Student_Count_Asian))
  crimes_by_area <- append(crimes_by_area, count(df_short$`Community Area` == area))
}
# for (temp in temps) {
#   crimes_by_temp <- append(crimes_by_temp, count(df_short$temp == temp))
# }
print()
```


```{r}
crimes_asians_frame <- data.frame(asians_by_area, crimes_by_area)
log_asians = log(asians_by_area)
log_crimes = log(crimes_by_area)
asian_model <- lm(log_crimes ~ log_asians, crimes_asians_frame)
print(summary(asians_model))
summary(asian_model)
```
```{r}
crimes_by_temp_frame <- data.frame(temps, crimes_by_temp)
ggplot(crimes_by_temp_frame, aes(x = temps, y = crimes_by_temp)) +
    geom_point() +
    geom_smooth(method = "lm")
temp_model <- lm(crimes_by_temp ~ temps, crimes_by_temp_frame)
print(summary(temp_model))
ggplot(crimes_asians_frame, aes(x = log_asians, y = log_crimes)) +
  geom_point()
```

```{r}
```

```{r}
table(df_short)
head(df_short)
summary(df_short)
ggplot(df_short, aes(x = Date)) +
  geom_histogram()
histogram(~date|`Community Area`, data=df_short,
          type="density",
          xlab="Date",
          main="Crimes Over Time by Community Area")
          # breaks=seq(from=60,to=190,by=10),
          # layout=c(1,2))
# ggplot(df_short, aes(x = Date, y=`Community Area`)) +
#   geom_point()
```
